---
title: "`srvyr` compared to the `survey` package"
author: "Greg Freedman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{srvyr Compared to the survey Package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `srvyr` package aims to add `dplyr` like syntax to the `survey` package.
This vignette focuses on how `srvyr` compares to the `survey` package, for more
information about survey design and analysis, check out the vignettes in the `survey`
package, or Thomas Lumley's book, [*Complex Surveys: A Guide to Analysis
Using R*](http://r-survey.r-forge.r-project.org/svybook/).

Everything that `srvyr` can do, can also be done in `survey`. In fact, behind
the scenes the `survey` package is doing all of the hard work for `srvyr`. 
`srvyr` strives to make your code simpler and more easily readable to you, 
especially if you are already used to the `dplyr` package.

# Motivating example

The `dplyr` package has made it fast and easy to write code to summarize data.
For example, if we wanted to check how the year-to-year change in academic 
progress indicator score varied by school level and percent of parents were 
high school graduates, we can do this:

```{r, message = FALSE, fig.width = 6}
library(survey)
library(ggplot2)
library(dplyr)

data(api)

out <- apistrat %>%
  mutate(hs_grad_pct = cut(hsg, c(0, 20, 100), include.lowest = TRUE,
                           labels = c("<20%", "20+%"))) %>%
  group_by(stype, hs_grad_pct) %>%
  summarize(api_diff = weighted.mean(api00 - api99, pw),
            n = n())

ggplot(data = out, aes(x = stype, y = api_diff, group = hs_grad_pct, fill = hs_grad_pct)) +
  geom_col(stat = "identity", position = "dodge") +
  geom_text(aes(y = 0, label = n), position = position_dodge(width = 0.9), vjust = -1)
```

However, if we wanted to add error bars to the graph to capture the uncertainty due to sampling
variation, we have to completely rewrite the `dplyr` code for the `survey` package.
`srvyr` allows a more direct translation.

# Preparing a survey dataset

`as_survey_design()`, `as_survey_rep()` and `as_survey_twophase()` are analogous
to `survey::svydesign()`, `survey::svrepdesign()` and `survey::twophase()` 
respectively. Because they are designed to match `dplyr`'s style of non-standard
evaluation, they accept bare column names instead of formulas (~). They also
move the data argument first, so that it is easier to use `magrittr` pipes
(`%>%`).

```{r, message = FALSE}
library(srvyr)

# simple random sample
srs_design_srvyr <- apisrs %>% as_survey_design(ids = 1, fpc = fpc)

srs_design_survey <- svydesign(ids = ~1, fpc = ~fpc, data = apisrs)
```

The `srvyr` functions also accept `dplyr::select()`'s special selection
functions (such as `starts_with()`, `one_of()`, etc.), so these functions are
analogous:
```{r, message = FALSE}
# selecting variables to keep in the survey object (stratified example)
strat_design_srvyr <- apistrat %>%
  as_survey_design(1, strata = stype, fpc = fpc, weight = pw,
                variables = c(stype, starts_with("api")))

strat_design_survey <- svydesign(~1, strata = ~stype, fpc = ~fpc,
                                 variables = ~stype + api99 + api00 + api.stu,
                                 weight = ~pw, data = apistrat)
```

The function `as_survey()` will automatically choose between the three `as_survey_*` 
functions based on the arguments, so you can save a few keystrokes.

```{r, message = FALSE}
# simple random sample (again)
srs_design_srvyr2 <- apisrs %>% as_survey(ids = 1, fpc = fpc)
```

# Data manipulation
Once you've set up your survey data, you can use `dplyr` verbs such as `mutate()`,
`select()`, `filter()` and `rename()`.

```{r, message = FALSE}
strat_design_srvyr <- strat_design_srvyr %>%
  mutate(api_diff = api00 - api99) %>%
  rename(api_students = api.stu)

strat_design_survey$variables$api_diff <- strat_design_survey$variables$api00 -
  strat_design_survey$variables$api99
names(strat_design_survey$variables)[names(strat_design_survey$variables) == "api.stu"] <- "api_students"
```

Note that `arrange()` is not available, because the `srvyr` object expects to
stay in the same order. Nor are two-table verbs such as `full_join()`,
`bind_rows()`, etc. available to `srvyr` objects either because they may have
implications on the survey design. If you need to use these functions, you
should use them earlier in your analysis pipeline, when the objects are still
stored as `data.frame`s.

# Summary statistics

## Of the entire population
`srvyr` also provides `summarize()` and several survey-specific functions that
calculate summary statistics on numeric variables: `survey_mean()`, `survey_total()`,
`survey_quantile()` and `survey_ratio()`. These functions differ from their
counterparts in `survey` because they always return a data.frame in a consistent
format. As such, they do not return the variance-covariance matrix, and so are
not as flexible.

```{r, message = FALSE}
# Using srvyr
out <- strat_design_srvyr %>%
  summarize(api_diff = survey_mean(api_diff, vartype = "ci"))

out

# Using survey
out <- svymean(~api_diff, strat_design_survey)

out
confint(out)
```

## By group

`srvyr` also allows you to calculate statistics on numeric variables by group, 
using `group_by()`.
```{r, message = FALSE}
# Using srvyr
strat_design_srvyr %>%
  group_by(stype) %>%
  summarize(api_increase = survey_total(api_diff >= 0),
            api_decrease = survey_total(api_diff < 0))

# Using survey
svyby(~api_diff >= 0, ~stype, strat_design_survey, svytotal)
```

## Proportions by group

You can also calculate the proportion or count in each group of a factor 
or character variable by leaving x empty in `survey_mean()` or `survey_total()`.

```{r, message = FALSE}
# Using srvyr
srs_design_srvyr %>%
  group_by(awards) %>%
  summarize(proportion = survey_mean(),
            total = survey_total())

# Using survey
svymean(~awards, srs_design_survey)
svytotal(~awards, srs_design_survey)
```

## Unweighted calculations

Finally, the `unweighted()` function can act as an escape hatch to calculate unweighted
calculations on the dataset.

```{r, message = FALSE}
# Using srvyr
strat_design_srvyr %>%
  group_by(stype) %>%
  summarize(n = unweighted(n()))

# Using survey
svyby(~api99, ~stype, strat_design_survey, unwtd.count)
```

# Back to the example

So now, we have all the tools needed to create the first graph and add error bounds.
Notice that the data manipulation code is nearly identical to the `dplyr` code, with a
little extra set up, and replacing `weighted.mean()` with `survey_mean`. 

```{r, message = FALSE, fig.width = 6}
strat_design <- apistrat %>%
  as_survey_design(strata = stype, fpc = fpc, weight  = pw)

out <- strat_design %>%
  mutate(hs_grad_pct = cut(hsg, c(0, 20, 100), include.lowest = TRUE,
                           labels = c("<20%", "20+%"))) %>%
  group_by(stype, hs_grad_pct) %>%
  summarize(api_diff = survey_mean(api00 - api99, vartype = "ci"),
            n = unweighted(n()))

ggplot(data = out, aes(x = stype, y = api_diff, group = hs_grad_pct, fill = hs_grad_pct,
                       ymax = api_diff_upp, ymin = api_diff_low)) +
  geom_col(stat = "identity", position = "dodge") +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.1) +
  geom_text(aes(y = 0, label = n), position = position_dodge(width = 0.9), vjust = -1)


```


# Grab Bag

## Using `survey` functions on `srvyr` objects

Because `srvyr` objects are just `survey` objects with some extra structure,
all of the functions from `survey` will still work with them. If you need to calculate
something beyond simple summary statistics, you can use `survey` functions.

```{r, message = FALSE}
glm <- svyglm(api00 ~ ell + meals + mobility, design = strat_design)
summary(glm)
```


## Standard evaluation

Srvyr now supports the standard evaluation conventions introduced in dplyr version 0.7 and
rlang. If you'd like to use a function programmatically, you can use the functions from 
rlang like `rlang::quo()` or `rlang::sym()` to capture the expression and `rlang::!!`
to unquote it. All of these functions are re-exported by srvyr, so you don't need
to load the rlang library to use them.

Here's a quick example, but please see the dplyr vignette 
[`vignette("programming", package = "dplyr")`](https://dplyr.tidyverse.org/articles/programming.html)
for more details.

```{r, message = FALSE}
fpc_var <- sym("fpc")
srs_design_srvyr <- apisrs %>% as_survey_design(fpc = !!fpc_var)

grouping_var <- sym("stype")
api_diff <- quo(api00 - api99)

srs_design_srvyr %>%
  group_by(!!grouping_var) %>% 
  summarize(
    api_increase = survey_total((!!api_diff) >= 0),
    api_decrease = survey_total((!!api_diff) < 0)
  )


```

Srvyr will also follow dplyr's lead on deprecating the old method of NSE, the
so-called "underscore functions" (like `summarize_`). Currently, they have been
soft-deprecated, but I expect them to be removed altogether in some future version
of srvyr.

## Scoped functions
Srvyr has also been able to take advantage of the new-ish dplyr "scoped" variants of 
the main manipulation functions like `summarize_at()`. These functions allow you to 
quickly calculate summary statistics for multiple variables. For example:

```{r}
# Calculate survey mean for all variables that have names starting with "api"
strat_design %>%
  summarize_at(vars(starts_with("api")), survey_mean)
```

The dplyr documentation `dplyr::scoped` provides more details.
