---
title: "srvyr compared to the survey package"
author: "Greg Freedman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The `srvyr` package aims to add `dplyr` like syntax to the `survey` package.
This vignette focuses on how `srvyr` compares to the `survey` package, for more
information about survey design and analysis, check out the vignettes in the `survey`
package, or Thomas Lumley's book, [*Complex Surveys: A Guide to Analysis
Using R*](http://r-survey.r-forge.r-project.org/svybook/).

Everything that `srvyr` can do, can also be done in `survey`. In fact, behind the scenes
the `survey` package is doing all of the hard work for `srvyr`. 
`srvyr` strives to make your code simpler and more easily readable to you,
especially if you are already used to the `dplyr` package.

# Motivating example

The `dplyr` package has made it fast and easy to write code to summarize data.
For example, if we wanted to check how the year-to-year change in academic progress indicator score
varied by school level and percent of parents were high school graduates, we can do this:

```{r, message = FALSE, fig.width = 6}
library(survey)
library(ggplot2)
library(dplyr)

data(api)

out <- apistrat %>%
  mutate(hs_grad_pct = cut(hsg, c(0, 20, 100), include.lowest = TRUE,
                           labels = c("<20%", "20+%"))) %>%
  group_by(stype, hs_grad_pct) %>%
  summarize(api_diff = weighted.mean(api00 - api99, pw),
            n = n())

ggplot(data = out, aes(x = stype, y = api_diff, group = hs_grad_pct, fill = hs_grad_pct)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(y = 0, ymax = api_diff, label = n), position = position_dodge(width = 0.9), vjust = -1)
```

However, if we wanted to add error bars to the graph to capture the uncertainty due to sampling
variation, we have to completely rewrite the `dplyr` code for the `survey` package.
`srvyr` allows a more direct translation.

# Preparing a survey dataset

`design_survey()` and `design_survey_rep()` are analagous to `survey::svydesign()`
and `survey::svrepdesign()` respectively. Because they are designed to match
`dplyr`'s style of non-standard evaluation, they accept bare column names instead
of formulas (~). They also move the data argument first, so that it is easier to use
`magrittr` pipes (`%>%`).

```{r, message = FALSE}
library(srvyr)

# simple random sample
srs_design_srvyr <- apisrs %>% design_survey(ids = 1, fpc = fpc)

srs_design_survey <- svydesign(ids = ~1, fpc = ~fpc, data = apisrs)
```

The `srvyr` functions also accept `dplyr::select()`'s special selection
functions (such as `starts_with()`, `one_of()`, etc.), so these functions are
analagous:
```{r, message = FALSE}
# selecting variables to keep in the survey object (stratified example)
strat_design_srvyr <- apistrat %>%
  design_survey(1, strata = stype, fpc = fpc, weight = pw,
                variables = c(stype, starts_with("api")))

strat_design_survey <- svydesign(~1, strata = ~stype, fpc = ~fpc,
                                 variables = ~stype + api99 + api00 + api.stu,
                                 weight = ~pw, data = apistrat)
```

# Data manipulation
Once you've set up your survey data, you can use `dplyr` verbs such as `mutate()`,
`select()`, `filter()` and `rename()`.

```{r, message = FALSE}
strat_design_srvyr <- strat_design_srvyr %>%
  mutate(api_diff = api00 - api99) %>%
  rename(api_students = api.stu)

strat_design_survey$variables$api_diff <- strat_design_survey$variables$api00 -
  strat_design_survey$variables$api99
names(strat_design_survey$variables)[names(strat_design_survey$variables) == "api.stu"] <- "api_students"
```

Note that `arrange()` is not available, because the `srvyr` object expects to stay in
the same order. Nor are two-table verbs such as `full_join()`, `bind_rows()`, etc.
available to `srvyr` objects either because they may have implications on the
survey design. If you need to use these functions, you should use them earlier in your
analysis pipeline, when the objects are still stored as `data.frame`s.

# Summary statistics

## Of the entire population
`srvyr` also provides `summarize()` and several survey-specific functions that calculate
summary statistics: `survey_mean()`, `survey_total()`, `survey_quantile()` and
`survey_ratio()`. These functions differ from their counterparts in `survey` because they
always return a data.frame in a consistent format. As such, they do not return the
variance-covariance matrix, and so are not as flexible.

```{r, message = FALSE}
# Using srvyr
out <- strat_design_srvyr %>%
  summarize(api_diff = survey_mean(api_diff, vartype = "ci"))

out

# Using survey
out <- svymean(~api_diff, strat_design_survey)

out
confint(out)
```

## By group

`srvyr` also allows you to calculate statistics by group, using `group_by()`.
```{r, message = FALSE}
# Using srvyr
strat_design_srvyr %>%
  group_by(stype) %>%
  summarize(api_increase = survey_total(api_diff >= 0),
            api_decrease = survey_total(api_diff < 0))

# Using survey
svyby(~api_diff >= 0, ~stype, strat_design_survey, svytotal)
```

## Unweighted calculations

Finally, the `unweighted()` function can act as an escape hatch to calculate unweighted
calculations on the dataset.

```{r, message = FALSE}
# Using srvyr
strat_design_srvyr %>%
  group_by(stype) %>%
  summarize(n = unweighted(n()))

# Using survey
svyby(~api99, ~stype, strat_design_survey, unwtd.count)
```

# Back to the example

So now, we have all the tools needed to create the first graph and add error bounds.
Notice that the data manipulation code is nearly identical to the `dplyr` code, with a
little extra set up, and replacing `weighted.mean()` with `survey_mean`. 

  ```{r, message = FALSE, fig.width = 6}
strat_design <- apistrat %>%
  design_survey(strata = stype, fpc = fpc, weight  = pw)

out <- strat_design %>%
  mutate(hs_grad_pct = cut(hsg, c(0, 20, 100), include.lowest = TRUE,
                           labels = c("<20%", "20+%"))) %>%
  group_by(stype, hs_grad_pct) %>%
  summarize(api_diff = survey_mean(api00 - api99, vartype = "ci"),
            n = unweighted(n()))

ggplot(data = out, aes(x = stype, y = api_diff, group = hs_grad_pct, fill = hs_grad_pct,
                       ymax = api_diff_upp, ymin = api_diff_low)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(position = position_dodge(width = 0.9), width = 0.1) +
  geom_text(aes(y = 0, label = n), position = position_dodge(width = 0.9), vjust = -1)


```

# Extras

## Using `srvyr` objects with `survey` functions

Because `srvyr` objects are just `survey` objects with some extra structure,
all of the functions from `survey` will still work with them. If you need to calculate
something beyond simple summary statistics, you can use `survey` functions.

```{r, message = FALSE}
glm <- svyglm(api00 ~ ell + meals + mobility, design = strat_design)
summary(glm)
```


## Standard evaluation

Similar to `dplyr`, the functions in `srvyr` have standard evaluation counterparts
with a "_" at the end. These functions can be useful when programming because they
make it easier to send arguments from other functions.

```{r, message = FALSE}
srs_design_srvyr <- apisrs %>% design_survey_(ids = "1", fpc = "fpc")

strat_design_srvyr %>%
  group_by_("stype") %>%
  summarize_(api_increase = "survey_total(api_diff >= 0)",
             api_decrease = "survey_total(api_diff < 0)")


```

For more information, see the standard evaluation vignette in `dplyr`, or `lazyeval`'s
vignettes.
