---
title: "Databases in srvyr"
author: "Greg Freedman Ellis"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{databases-srvyr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, echo = FALSE, message = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Srvyr now has experimental support of database backed surveys. It uses dplyr databases 
as the backend, not the survey package's database support, which allows for nearly identical 
consistent syntax when working with database or local data.frame backed surveys. 
The actual implementation is similar to the survey package's in that it stores the 
survey variables (such as weighting, stratum, etc.) in local memory and so will provide the most
benefit when dealing with surveys with a large number of variables, rather than surveys
with a large number of observations. It is slower than the survey package's support, because it is
more careful about the sorting order when pulling data from the database, in my testing 
it is about 2X slower than survey. 

During development, I have tested using SQLite and MonetDBLite database, but in theory 
other database backends should work as well.

This vignette shows the basics of how to use srvyr with databases. 
It is based on analysis from the wonderful resource asdfree 
( [website](http://asdfree.com/) and [github](https://github.com/ajdamico/asdfree) ). 
Many thanks to [ajdamico](http://github.com/ajdamico) and collaborators. 
Specifically, I have adpated code from [American Community Survey - 2011 single year analysis](https://github.com/ajdamico/asdfree/blob/d0b965e9672161da086e30471760d08281b74343/American%20Community%20Survey/2011%20single-year%20-%20analysis%20examples.R) and the associated data preparation scripts. 

## Database Setup ##
In order to focus on srvyr and databases, we start with a prepared dataset. The full code 
is avaliable on Github TKTKTKT, and the high level description of what it does is:

- Download data from acs website (currently only Alaska and Hawaii to save time, 
though it would be easy to adapt to download to all 50 states and Puerto Rico). 

- Merges the household and person datasets so that we can look at the variables
related to each person including those at the household level

- Selects only a few variables that will be used in this analysis to save space, but
again it could easily be adapted to keep all of the variables. 


For more information on the specifics of the American Community Survey, see 
the asdfree site. Now, our code loads this prepared dataset, initiates a MonetDB database, and 
puts the data into the dataset.

```{r}
suppressMessages({
  library(MonetDBLite)
  library(survey)
  library(srvyr)
  library(dplyr)
  library(RSQLite)
})

# Load data (Currently only Alaska and Hawaii to keep file size small and with 
# limited variables, butcode that downloaded the files is available here TKTKTKT,
# and could easily be adapted to download all states.)
load("acs_m.Rdata") # acs_m data

# Set up database and table
db <- src_monetdblite()
acs_m_db <- copy_to(db, acs_m, "acs_m", temporary = FALSE)

# Make table read only to reflect real world usage (see advanced topics below).
db_status <- dbSendQuery(acs_m_db$src$con, "ALTER TABLE acs_m SET READ ONLY")

# Or, if the data was already stored in the database, you could do this
# acs_m_data <- tbl(db, sql("SELECT * FROM acs_m")) 
```

Now that we have the data in the database, we can interact with the database directly using sql 
commands, or we can use dplyr's functionality to treat it mostly the same as a local `data.frame`.
However, the data is not stored in memory, so we could work with much larger datasets (though
in this case, the data is too small for this to be a problem).

```{r}
# Same results
acs_m %>% 
  group_by(sex) %>%
  summarize(hicov = mean(hicov))

acs_m_db %>% 
  group_by(sex) %>%
  summarize(hicov = mean(hicov))

# But smaller object size
object.size(acs_m)
object.size(acs_m_db)

```

Note that though many commands behave exactly the same whether on a local data.frame or
database, sometimes more advanced / complicated syntax around variable modification allowed 
in dplyr does not work on a particular database and so it is better to be more explicit. For
example, creating a variable inside of a summarize call does not work (at least with MonetDB).
.

```{r}
acs_m %>% 
  group_by(sex) %>%
  summarize(hicov = mean(hicov == 1))

# acs_m_db %>% 
#   group_by(sex) %>%
#   summarize(hicov = mean(hicov == 1))
#
# > Error in .local(conn, statement, ...) :
# > Unable to execute statement....
# > ....

# Creating the variable separately works as an integer works though
acs_m_db %>% 
  group_by(sex) %>%
  mutate(hicov = ifelse(hicov == 1, 1L, 0L)) %>%
  summarize(hicov = mean(hicov))
```

Further, sometimes working with variable types can get difficult if you are used to working in
R. Notice how in the above, instead of `hicov = (hicov == 1)`, I wrote out the ifelse statement. 
If I hadn't MonetDB would be unable to calculate the mean of the boolean variable created.

Finally, a major difference when transitioning from dplyr on local data.frames is that not 
all R functions are translated to SQL. For example, `cut()` isn't implemented in SQL, 
so you can't create a new variable in the data.frame using it. 

```{r}
acs_m %>% 
  group_by(agecat = cut(agep, c(0, 19, 35, 50, 65, 200))) %>%
  summarize(hicov = mean(hicov == 1))

# acs_m_db %>% 
#   group_by(agecat = cut(agep, c(0, 19, 35, 50, 65, 200))) %>%
#   summarize(hicov = mean(hicov == 1))
#
# > Error in .local(conn, statement, ...) :
# > Unable to execute statement....
# > ...

acs_m_db %>% 
  mutate(agecat = ifelse(agep < 19, "0-18", 
                         ifelse(agep >= 19 & agep < 35, "19-34", 
                                ifelse(agep >= 35 & agep < 50, "35-49", 
                                       ifelse(agep >= 50 & agep < 65, "50-64", 
                                              ifelse(agep >= 65, "65+", NA)))))) %>%
  group_by(agecat) %>% 
  summarize(hicov = mean(hicov))
```

For more information on the specifics of databases
with dplyr, see `vignette("database", package = dplyr)`, the `DBI` package or the 
specific database packages, like `MonetDBLite` or `RSQLite`.

## Srvyr Setup ##
Srvyr commands are nearly identical to old. The only difference for setup is that you 
need a variable that uniquely identifieds each row in the database (uid). 

```{r}
acs_m_db_svy <- acs_m_db %>% 
  as_survey_rep(
    weight = pwgtp,
    repweights = matches("pwgtp[0-9]+") ,
    scale = 4 / 80,
    rscales = rep(1 , 80),
    mse = TRUE,
    type = "JK1", 
    variables = -c(matches("^pwgtp")),
    uid = c(serialno, sporder)
  )

acs_m_db_svy
```

Because srvyr stores the survey variables locally, the srvyr object takes up
much more memory than the dplyr one. However, this object would not grow in size
if you added more data variables to your survey, so if your survey is very wide,
it will save a lot space.

```{r}
object.size(acs_m_db_svy)
```

# Analysis # 

Analysis commands from srvyr are also similar to ones that work on local data.frames. 
The main differences come from the issues discussed above about explicitly creating variables
difficulties in translating R commands, and variable types. 

The following anlaysis is based on the asdfree analysis and shows some basi analysis on
the total populaiton, insurance coverage, age and sex. 

```{r}
# You can calculate the population of the united states #
# by state
acs_m_db_svy %>%
  mutate(one = 1L) %>% # Note that because of weird behavior of MonetDB, need to use 1L not just 1
  group_by(st) %>% 
  summarize(count = survey_total(one))

# Or the average age of downloaded states
acs_m_db_svy %>%
  summarize(agep = survey_mean(agep, na.rm = TRUE))

# Average age by state
acs_m_db_svy %>%
  group_by(st) %>% 
  summarize(agep = survey_mean(agep, na.rm = TRUE))

# percent uninsured - nationwide (of downloaded states)
acs_m_db_svy %>%
  mutate(hicov = as.character(hicov)) %>% 
  group_by(hicov) %>% 
  summarize(pct = survey_mean(na.rm = TRUE))

# by state
acs_m_db_svy %>%
  mutate(hicov = as.character(hicov)) %>% 
  group_by(st, hicov) %>% 
  summarize(pct = survey_mean(na.rm = TRUE))


# 25th, median, and 75th percentile of age of residents of the united states (downloaded states)
acs_m_db_svy %>%
  summarize(agep = survey_quantile(agep, c(0.25, 0.5, 0.75), na.rm = TRUE))


# Filter works, so we can restrict the acs_m object to females only
acs_m_db_svy_female <- acs_m_db_svy %>%
  filter(sex == 2)

# Now any of the above commands can be re-run using the acs_m_female object
# instead of the acs_m object in order to analyze females only

# This is equivalent to using acs_m, and applying the filter every time.

# average age - nationwide (of downloaded states), restricted to females
acs_m_db_svy_female %>%
  summarize(agep = survey_mean(agep, na.rm = TRUE))

# median age - nationwide (of downloaded states), restricted to females
acs_m_db_svy_female %>%
  summarize(agep = survey_median(agep, na.rm = TRUE))


# Note that though some R functions are translated by dplyr into SQL, not
# all of them are. For example, when constructing a new age category 
# variable in the dataset neither findIntervals nor cut work on databases, 
# so we have to spell out groups with ifelse()
acs_m_db_svy %>% 
  mutate(agecat = ifelse(agep < 19, "0-18", 
                         ifelse(agep >= 19 & agep < 35, "19-34", 
                                ifelse(agep >= 35 & agep < 50, "35-49", 
                                       ifelse(agep >= 50 & agep < 65, "50-64", 
                                              ifelse(agep >= 65, "65+", NA)))))) %>%
  group_by(agecat) %>% 
  summarize(pct = survey_mean(na.rm = TRUE))
```

## Advanced Topics ##
### Write Access ###
Note that srvyr does not require write access to perform calculations, 
the database created in this vignette was set to read-only at the beginning. 
This can be important when you want to make sure that your original data is
not altered accidentally, or if you don't have write access to a database.

### Speed ###
Srvyr using a database is significantly slower than using the survey package's
database support, which is in turn slower than using local data.frames. The real
advantages of databases are for working with sureys that are very wide (many variables). 
Srvyr's advantage over the survey package comes down to the syntax, if you prefer the
consistent style of srvyr you'll need to be willing to stomach the performance loss.
