---
title: "srvr speed"
author: "Greg Freedman"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

This vignette is for testing only and excluded from the R build because it is too slow 
(~5 min on my machine).

# Scaling number of observations 

- Scales linearly with number of observations
- No major differences between survey and srvyr
- Using "calls" is slower than the variable itself in survey

```{r, message = FALSE}
library(survey)
library(srvyr)
library(dplyr)
library(microbenchmark)
library(tidyr)
library(ggplot2)
library(scales)

micro_times <- 5

set.seed(1)

num_obs <- c(1e4, 1e5, 1e6)
num_vars <- 10
num_strata <- 10


function_list <- list(
  mean = list(f1 = svymean, 
              f2 = svymean, 
              f3 = function(svy, x) summarize_each(svy, funs(survey_mean(.)), one_of(x))
  ), 
    total = list(f1 = svytotal, 
                 f2 = svytotal, 
                 f3 = function(svy, x) summarize_each(svy, funs(survey_total(.)), one_of(x))
  ), 
  median = list(f1 = function(x, svy) svyquantile(x, svy, quantiles = 0.5), 
                f2 = function(x, svy) svyquantile(x, svy, quantiles = 0.5), 
                f3 = function(svy, x) summarize_each(svy, funs(survey_median(.)), one_of(x))
  )
)


data <- lapply(num_obs, function(obs) {
  out <- data.frame(strata = sample(paste0("st", seq_len(num_strata)), obs, replace = TRUE), 
                       probs = runif(obs))
  out[, c(3:(2 + num_vars))] <- runif(obs * num_vars) + rep(seq_len(num_vars), each = obs)
  out
})


svy <- list()
srvyr <- list()
out_setup <- lapply(seq_along(data), function(ddd) {
  microbenchmark(`svy setup` = svy[[ddd]] <<- svydesign(~1, strata = ~strata, probs = ~probs, data = data[[ddd]]), 
                 `srvyr setup` = srvyr[[ddd]] <<- data[[ddd]] %>% as_survey_design(strata = strata, probs = probs), 
                 times = micro_times, unit = "s")
})

names(out_setup) <- as.character(num_obs)
out_setup <- out_setup %>% 
  bind_rows(.id = "obs") %>%
  extract(expr, c("expr", "fff"), "(.+) (.+)")

names(svy) <- as.character(num_obs)
names(srvyr) <- as.character(num_obs)

out <- lapply(function_list, function(fff) {
  aaa <- lapply(as.character(num_obs), function(ddd) {
    microbenchmark(`svy call` = fff$f1(~V3, svy[[ddd]]), 
                   `svy var` = fff$f2(data.frame(svy[[ddd]]$variables$V3), svy[[ddd]]),
                   `srvyr` = srvyr[[ddd]] %>% fff$f3("V3"), 
                   times = micro_times, unit = "s")
  })
  names(aaa) <- as.character(num_obs)
  aaa
})

out <- lapply(names(function_list), function(fff) {
  out[[fff]] %>% 
    bind_rows(.id = "obs") %>%
    mutate(fff = fff)
}) %>% 
  bind_rows() %>% 
  bind_rows(out_setup) %>% 
  group_by(obs, expr, fff) %>%
  summarize(time = mean(time)) %>% 
  ungroup() %>%
  mutate(obs = as.numeric(obs), 
         fff = factor(fff, c("setup", "mean", "total", "median")), 
         time = time / 1000000000)
  

ggplot(data = out, aes(x = obs, y = time, group = expr, color = expr)) + 
  geom_point() + geom_line() + 
  facet_wrap(~fff) + 
  scale_x_continuous(labels = comma)
  
```


# Scaling number of variables 

- srvyr is much worse than survey (scales linearly while survey scales sublinearly)
- median is way faster in srvyr - I'm not sure why...
- sending variables into survey is better than calls again

```{r, message = FALSE}
num_obs <- 1e5
num_vars <- c(1, 5, 10, 20)
num_strata <- 10

data <- lapply(num_vars, function(vars) {
  out <- data.frame(strata = sample(paste0("st", seq_len(num_strata)), num_obs, replace = TRUE), 
                       probs = runif(num_obs))
  out[, c(3:(2 + vars))] <- runif(num_obs * vars) + rep(seq_len(vars), each = num_obs)
  out
})


svy <- list()
srvyr <- list()
out_setup <- lapply(seq_along(data), function(ddd) {
  microbenchmark(`svy setup` = svy[[ddd]] <<- svydesign(~1, strata = ~strata, probs = ~probs, data = data[[ddd]]), 
                 `srvyr setup` = srvyr[[ddd]] <<- data[[ddd]] %>% as_survey_design(strata = strata, probs = probs), 
                 times = micro_times, unit = "s")
})

names(out_setup) <- as.character(num_vars)
out_setup <- out_setup %>% 
  bind_rows(.id = "vars") %>%
  extract(expr, c("expr", "fff"), "(.+) (.+)")

names(svy) <- as.character(num_vars)
names(srvyr) <- as.character(num_vars)

out <- lapply(function_list, function(fff) {
  aaa <- lapply(as.character(num_vars), function(ddd) {
    vnames <- names(svy[[ddd]]$variables)[grep("^V", names(svy[[ddd]]$variables))]
    microbenchmark(`svy call` = fff$f1(make.formula(vnames), svy[[ddd]]), 
                   `svy var` = fff$f2(data.frame(svy[[ddd]]$variables[vnames]), svy[[ddd]]),
                   `srvyr` = srvyr[[ddd]] %>% fff$f3(vnames), 
                   times = micro_times, unit = "s")
  })
  names(aaa) <- as.character(num_vars)
  aaa
})

out <- lapply(names(function_list), function(fff) {
  out[[fff]] %>% 
    bind_rows(.id = "vars") %>%
    mutate(fff = fff)
}) %>% 
  bind_rows() %>% 
  bind_rows(out_setup) %>% 
  group_by(vars, expr, fff) %>%
  summarize(time = mean(time)) %>% 
  ungroup() %>%
  mutate(vars = as.numeric(vars), 
         fff = factor(fff, c("setup", "mean", "total", "median")), 
         time = time / 1000000000)
  

ggplot(data = out, aes(x = vars, y = time, group = expr, color = expr)) + 
  geom_point() + geom_line() + 
  facet_wrap(~fff) + 
  scale_x_continuous(labels = comma)
  
```


# Scaling number of strata (and grouping variables)

- srvyr just as fast as survey in calculations. 
- Slight penalty for setup (group_by), but not noticable at this number of obs / groups 
  (or even 1e6 obs, run separately)
- calculations grow linearly 
- Can't use survey variables because of bug in svyby

```{r, message = FALSE}
function_list <- list(
  mean = list(f1 = function(x, svy) svyby(x, ~V3, svy, svymean),
              f2 = function(x, svy) svyby(x, data.frame(svy$variables$V3), svy, svymean), 
              f3 = function(svy, x) summarize_each(group_by(svy, V3), funs(survey_mean(.)), one_of(x))
  ), 
  total = list(f1 = function(x, svy) svyby(x, ~V3, svy, svytotal),
               f2 = function(x, svy) svyby(x, data.frame(svy$variables$V3), svy, svytotal), 
               f3 = function(svy, x) summarize_each(group_by(svy, V3), funs(survey_total(.)), one_of(x))
  ), 
  median = list(f1 = function(x, svy) svyby(x, ~V3, svy, svyquantile, quantiles = 0.5, ci = TRUE),
                f2 = function(x, svy) svyby(x, data.frame(svy$variables$V3), svy, svyquantile, 
                                            quantiles = 0.5, ci = TRUE), 
                f3 = function(svy, x) summarize_each(group_by(svy, V3), funs(survey_median(.)), one_of(x))
  )
)


num_obs <- 1e5
num_vars <- 2
num_strata <- 10
num_groups <- c(2, 5, 10, 20)

data <- lapply(num_groups, function(group) {
  out <- data.frame(strata = sample(paste0("st", seq_len(num_strata)), num_obs, replace = TRUE), 
                       probs = runif(num_obs))
  out[, c(3:(2 + num_vars))] <- runif(num_obs * num_vars) + rep(seq_len(num_vars), each = num_obs)
  out$V3 <- cut_interval(out$V3, group)
  out
})

svy <- list()
srvyr <- list()
out_setup <- lapply(seq_along(data), function(ddd) {
  microbenchmark(`svy setup` = svy[[ddd]] <<- svydesign(~1, strata = ~strata, probs = ~probs, data = data[[ddd]]), 
                 `srvyr setup` = srvyr[[ddd]] <<- data[[ddd]] %>% as_survey_design(strata = strata, probs = probs), 
                 times = micro_times, unit = "s")
})

names(out_setup) <- as.character(num_groups)
out_setup <- out_setup %>% 
  bind_rows(.id = "groups") %>%
  extract(expr, c("expr", "fff"), "(.+) (.+)")

names(svy) <- as.character(num_groups)
names(srvyr) <- as.character(num_groups)

out <- lapply(function_list, function(fff) {
  aaa <- lapply(as.character(num_groups), function(ddd) {
    vnames <- "V4"
    suppressWarnings(microbenchmark(`svy call` = fff$f1(make.formula(vnames), svy[[ddd]]), 
                   #`svy var` = fff$f2(vnames, svy[[ddd]]), # doesn't work currently bc of bug in svyby
                   `srvyr` = srvyr[[ddd]] %>% fff$f3(vnames), 
                   times = micro_times, unit = "s"))
  })
  names(aaa) <- as.character(num_groups)
  aaa
})

out <- lapply(names(function_list), function(fff) {
  out[[fff]] %>% 
    bind_rows(.id = "groups") %>%
    mutate(fff = fff)
}) %>% 
  bind_rows() %>% 
  bind_rows(out_setup) %>% 
  group_by(groups, expr, fff) %>%
  summarize(time = mean(time)) %>% 
  ungroup() %>%
  mutate(groups = as.numeric(groups), 
         fff = factor(fff, c("setup", "mean", "total", "median")), 
         time = time / 1000000000)
  

ggplot(data = out, aes(x = groups, y = time, group = expr, color = expr)) + 
  geom_point() + geom_line() + 
  facet_wrap(~fff) + 
  scale_x_continuous(labels = comma)
  
```
